Higgs Assisted Razor Search for Higgsinos at 100 TeV

### 2019-07-23 main.tex
Added a couple of sentences about compressed spectra

\newcommand{\Shufang}[1]{{\bf\color{Maroon}  #1}}
\newcommand{\Adarsh}[1]{{\bf\color{RoyalBlue} AP: #1}}
\newcommand{\ap}[1]{\textcolor{RoyalBlue}{#1}}
\renewcommand{\H}{\widetilde{H}^0}
\newcommand{\B}{\widetilde{B}^0}
\newcommand{\N}{\widetilde{\chi}^0}
\newcommand{\image}[2]{\parbox{#1\textwidth}{\includegraphics[width=#1\textwidth]{images/#2.pdf}}}
\newcommand{\met}{{\slashed{E}_T}}

@@

 % Why go beyond the MSSM?
Though a natural MSSM spectrum has the potential to tame the hierarchy
problem, it is under siege from recent data from the Large Hadron
Collider (LHC) \cite{Aaboud:2018ujj, Sirunyan:2018vjp}. 
A compelling alternative scenario comes in the form of
split supersymmetry (split SUSY) \cite{Wells:2003tf,
ArkaniHamed:2004yi, Giudice:2004tc}. In this scenario, the lightest
superpartners are the fermionic ones, on the
scale of 1-10 TeV, while the scalar superpartners can be much heavier,
on the scale of 100 - 1000 TeV. In exchange for accepting some level of
fine-tuning, we obtain numerous benefits, including the suppression of
flavor-changing neutral currents and greater compatibility with data
from CP-violation experiments.  The low lying 
fermionic superpartners consist of neutralinos and charginos (collectively
known as electroweakinos), and gluinos. The current LHC search limits on
gluinos are already around 2 TeV for prompt decay and 2.5 TeV for disappearing track searches~\cite{CMS-PAS-SUS-19-005}.   The
limits on electroweakinos with heavy scalar superparticles,
however, are still relatively weak: around 650 $-$ 750 GeV or  less~\cite{ATL-PHYS-PUB-2019-022, Sirunyan:2017lae,Aaboud:2018htj,Sirunyan:2018ubx,Aaboud:2018zeb}, depending on
the decay modes of the parent electroweakinos and their species.  In this paper, we will
focus on the discovery potential of the neutralinos and charginos at the 100
TeV $pp$ collider.

@@

Since the Bino is the only supersymmetric particle lighter than the Higgsino in
the mass hierarchy we consider, the neutral Higgsino will decay to a Bino and a
neutral SM boson (\emph{Z/h}) with a branching ratio of 100\%.  Searches for
Higgsino NLPs with a Bino LSP  via  multi-lepton final states have been
studied in Ref.~\cite{Gori:2014oua} at a 100 TeV $pp$ collider, for pair production of $\widetilde\chi_{2,3}^0\widetilde\chi_1^\pm$ and  $\widetilde\chi_{1}^+\widetilde\chi_1^-$.    In this paper, we attempt to exploit the relatively high branching
ratio of the SM Higgs boson to \emph{b} quarks, and study the potential of the channel of
$\chi_2^0\chi_3^0\rightarrow Zh \chi_1^0\chi_1^0\rightarrow bb\ell\ell \met$.

@@

\begin{equation}
\mathbf{M}_{\widetilde{N}}=
\begin{pmatrix}
  M_1 & 0 & -\cb\sw\mz & \sinb\sw\mz \\
  0 & M_2 & \cb\cw\mz & -\sinb\cw\mz \\
  -\cb\sw\mz & \cb\cw\mz & 0 & -\mu \\
  \sinb\sw\mz & -\sinb\cw\mz & -\mu & 0
\end{pmatrix},
\end{equation}
where we employ the notation $s_W, c_W = \sin\theta_W, \cos\theta_W$
for $\theta_W$ being the Weinberg mixing angle. $s_\beta, c_\beta$ are defined
accordingly with $\tan\beta= v_u/v_d$, where $v_u$ and $v_d$ are the vacuum
expectation values of $H_u^0$ and $H_d^0$, with $v_u^2+v_d^2=v^2=(246\ {\rm
GeV})^2$. The parameters $M_1$, $M_2$, and $\mu$ are the mass parameters of
Bino, Wino, and Higgsinos.  The mass eigenstates are usually labeled as
$\chi_i^0$ for $i=1 \ldots 4$ with increasing mass eigenvalues.

@@

The optimal search strategy for finding electroweakinos   highly depends on
  their mass spectra \cite{Han:2013kza}.  Phenomenological collider studies on finding electroweakinos with a small mass splitting of about 0.1 - 50 GeV at a 100 TeV collider can be
found in \cite{Low:2014cba, Bramante:2014tba, Berlin:2015aba, Cirelli:2014dsa},
\Shufang{Say a few words about typical search strategy there.}
\Adarsh{Yes, I focused more on well separated spectra since that is more similar to our case - should I add more details for the 0.1 - 50 GeV compressed spectra studies?} \Shufang{Add a couple sentences here of the search strategy for the small mass splitting case.}
Well-separated spectra have been studied in \cite{Gori:2014oua,
Acharya:2014pua} using
multi-lepton channels, of the form 
\begin{align}
  pp\rightarrow \widetilde{\chi}_i\widetilde{\chi}_j\rightarrow VV +
  \widetilde{\chi}_1^0\widetilde{\chi}_1^0\rightarrow 2\ell/3\ell +  
  \met,
\end{align}
where $\widetilde{\chi}_{i,j}$ can be a neutralino or chargino,
$\widetilde{\chi}_1^0$ is the lightest neutralino, and $V$ being any of the SM gauge bosons $W$, $Z$ or scalar $h$.  These searches work well since
the large mass difference between the electroweakinos can lead to energetic
leptons that can be easily identified. Among the multi-lepton searches for
Higgsinos, the trilepton searches: $\widetilde\chi_{2,3}^0\widetilde\chi_1^\pm\rightarrow WZ\widetilde\chi_1^0\widetilde\chi_1^0\rightarrow \ell\ell\ell + \met$, with $\ell=e, \ \mu$,  have the best reach, due to the relatively high production cross-section of
chargino-neutralino pairs combined with the large reduction in $t\overline{t}$
and QCD backgrounds obtained by requiring three leptons.  Higgsino NLSPs can be excluded at 95\% C.L. with mass up to 2.8 TeV or be discovered with  mass up to 1.8 TeV, for a massless Bino LSP.   The reach reduces for larger $M_1$, which vanishes for $M_1\gtrsim 600$ GeV (1000 GeV) for discovery (exclusion).    Multilepton searches
with \emph{Zh} and \emph{ZZ} as the intermediate dibosons are unlikely to be as powerful,
due to the smaller pair-production cross-section of neutral Higgsinos, combined
with the low branching ratio of \emph{Z/h} to leptons. However, if we move beyond
multilepton searches, the channel with \emph{Zh} (originally suggested in
\cite{Han:2013kza}) as the intermediate dibosons emerges as a possible
competitor to the \emph{WZ} channel.

In the left panel of Fig.~\ref{fig:xsection_plot}, we show the pair-production cross section $\widetilde\chi_2^0\widetilde\chi_3^0$ for the Higgsino-like NLSPs.  The cross section varies between
200 fb to 0.5 fb for Higgsino masses between 500 GeV and 2500 GeV. The right
panel of Fig.~\ref{fig:xsection_plot} shows the decay branching fractions of
$\chi_{2,3}^0\rightarrow Z/h \chi_1^0$, as calculated in SUSY-HIT \cite{Djouadi:2006bz}.
While one of the Higgsinos mostly decays to $Z$, the other
Higgsino mostly decays to $h$.    Therefore,  for pair produced Higgsinos $\widetilde\chi_2^0\widetilde\chi_3^0$, the \emph{Zh} mode has the greatest
branching fraction.  After multiplying all the branching fractions, the signal cross sections of 
\begin{align}
pp\rightarrow \widetilde{\chi}_{2}^{0}\widetilde{\chi}_{3}^{0}\rightarrow
(Z\widetilde{\chi}_1^{0})(h\widetilde{\chi}_1^{0})\rightarrow
bb\ell\ell \chi_1^0\chi_1^0.
\end{align}
become comparable  to trilepton processes, as shown in    table \ref{tab:xsections}.   $Zh+\met$ signal thus provide an alternative
discovery channel for Higgsinos.   The signal contains two
\emph{b}-jets, two same flavor, opposite sign leptons, and missing transverse
energy. The main backgrounds for this process are: $t\overline{t}$, \emph{tbW} with
the $b$ and $W$ not coming from a \emph{t}, and \emph{bbWW} with no intermediate
$s$-channel top quarks.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.49\textwidth]{images/xsection_plot.pdf}
  \includegraphics[width=0.49\textwidth]{images/br_plot.pdf}
  \caption{Higgsino pair production $\widetilde\chi_2^0\widetilde\chi_3^0$
    cross section (left panel), calculated using
    Prospino2~\cite{Beenakker:1999xh},  and decay branching fractions
    (right panel), calculated in SUSY-HIT~\cite{Djouadi:2006bz},  as a function
    of $\mu$ for $M_1=25$ GeV. However, it should be noted that the dependence
    on $M_1$ is very weak.  \Shufang{Could you use $\mu$ for both xlabels?  The CS and Br will depend slightly on the sign of mu.  Since we do use positive mu in our analyses, let's label it this way.  I have a footnote in intro to explain the sign effect of mu.  Also, change right panel, ylabel to be either Br or Branching Fraction.}
}
  \label{fig:xsection_plot}
\end{figure}

\begin{table}
  \centering
  \begin{tabular}{l|rr}
    \toprule
    Stage & $\N_{2,3}\widetilde{\chi}^\pm_{1}$ & $\N_2\N_3$\\
    \midrule
    Pair production cross section & 60 fb & 17.2 fb\\
    Intermediate diboson contribution & ($WZ$) 29.7 fb  & ($Zh$) 9.15 fb \\
    Applying \emph{BR}$(W\rightarrow \ell\nu)$, \emph{BR}$(Z\rightarrow \ell\ell)$ \& \emph{BR}$(h\rightarrow bb)$ & 0.42 fb & 0.37 fb\\
    \bottomrule
  \end{tabular}
  \caption{Comparison of Higgsino pair production cross-sections for
    $(\widetilde{\chi}_{2,3}^0\widetilde{\chi}^\pm_{1})$ and
    $(\N_{2}\N_{3})$, for $\mu\approx 1$ TeV, at a 100
    TeV \emph{pp} collider. 
    The pair-production cross section for $\N_{2,3}\widetilde{\chi}^\pm_1$ is taken from
    \cite{Gori:2014oua}, while the pair-production cross-section for $\N_2\N_3$ is calculated at Next-to-Leading Order (NLO) using Prospino2. The branching ratios to the diboson intermediate states are calculated using SUSY-HIT, and the branching ratios to the SM final states are taken from
    \cite{Olive:2016xmw}.    }
  \label{tab:xsections}
\end{table}

>>>>

\newcommand{\Shufang}[1]{{\bf\color{Maroon}  #1}}
\newcommand{\Adarsh}[1]{{\bf\color{RoyalBlue} AP: #1}}
\newcommand{\ap}[1]{\textcolor{RoyalBlue}{#1}}
\renewcommand{\H}{\widetilde{H}^0}
\newcommand{\B}{\widetilde{B}^0}
\newcommand{\N}{\widetilde{\chi}^0}
\newcommand{\C}{\widetilde{\chi}^\pm}
\newcommand{\image}[2]{\parbox{#1\textwidth}{\includegraphics[width=#1\textwidth]{images/#2.pdf}}}
\newcommand{\met}{{\slashed{E}_T}}

@@

 % Why go beyond the MSSM?
Though a natural MSSM spectrum has the potential to tame the hierarchy problem,
it is under siege from recent data from the Large Hadron Collider (LHC)
\cite{Aaboud:2018ujj, Sirunyan:2018vjp}.  A compelling alternative scenario
comes in the form of split supersymmetry (split SUSY) \cite{Wells:2003tf,
ArkaniHamed:2004yi, Giudice:2004tc}. In this scenario, the lightest
superpartners are the fermionic ones, on the scale of 1-10 TeV, while the
scalar superpartners can be much heavier, on the scale of 100 - 1000 TeV. In
exchange for accepting some level of fine-tuning, we obtain numerous benefits,
including the suppression of flavor-changing neutral currents and greater
compatibility with data from CP-violation experiments.  The low lying fermionic
superpartners consist of neutralinos and charginos (collectively known as
electroweakinos), and gluinos. The current LHC search limits on gluinos are
already around 2 TeV for prompt decay and 2.5 TeV for disappearing track
searches~\cite{CMS-PAS-SUS-19-005}.   The limits on electroweakinos with heavy
scalar superparticles, however, are still relatively weak: around 650 $-$ 750
GeV or less~\cite{ATL-PHYS-PUB-2019-022, Sirunyan:2017lae, Aaboud:2018htj,
Sirunyan:2018ubx, Aaboud:2018zeb}, depending on the decay modes of the parent
electroweakinos and their species.  In this paper, we will focus on the
discovery potential of the neutralinos and charginos at the 100 TeV $pp$
collider.

@@

Since the Bino is the only supersymmetric particle lighter than the Higgsino in
the mass hierarchy we consider, the neutral Higgsino will decay to a Bino and a
neutral SM boson (\emph{Z/h}) with a branching ratio of 100\%.  Searches for
Higgsino NLPs with a Bino LSP  via  multi-lepton final states have been studied
in Ref.~\cite{Gori:2014oua} at a 100 TeV $pp$ collider, for pair production of
$\N_{2,3}\C_1$ and
$\widetilde\chi_{1}^+\widetilde\chi_1^-$. In this paper, we attempt to exploit
the relatively high branching ratio of the SM Higgs boson to \emph{b} quarks,
and study the potential of the channel of $\N_2\N_3\rightarrow Zh
\N_1\N_1\rightarrow bb\ell\ell \met$.

@@

\begin{align}
  \mathbf{M}_{\widetilde{N}}=
  \begin{pmatrix}
    M_1         & 0            & -\cb\sw\mz & \sinb\sw\mz \\
    0           & M_2          & \cb\cw\mz  & -\sinb\cw\mz \\
    -\cb\sw\mz  & \cb\cw\mz    & 0          & -\mu \\
    \sinb\sw\mz & -\sinb\cw\mz & -\mu       & 0
  \end{pmatrix},
\end{align}

\noindent where we employ the notation $s_W, c_W = \sin\theta_W, \cos\theta_W$
for $\theta_W$ being the Weinberg mixing angle. $s_\beta, c_\beta$ are defined
accordingly with $\tan\beta= v_u/v_d$, where $v_u$ and $v_d$ are the vacuum
expectation values of $H_u^0$ and $H_d^0$, with $v_u^2+v_d^2=v^2=(246\ {\rm
GeV})^2$. The parameters $M_1$, $M_2$, and $\mu$ are the mass parameters of
Bino, Wino, and Higgsinos.  The mass eigenstates are usually labeled as
$\chi_i^0$ for $i=1 \ldots 4$ with increasing mass eigenvalues.

@@

The optimal search strategy for finding electroweakinos   highly depends on
their mass spectra \cite{Han:2013kza}.  Phenomenological collider studies on
finding electroweakinos with a small mass splitting of about 0.1 - 50 GeV at a
100 TeV collider can be found in \cite{Low:2014cba, Bramante:2014tba,
Berlin:2015aba, Cirelli:2014dsa}. The projected exclusion reaches for pure Wino
LSPs (with a mass splitting of 164.6 MeV \cite{Ibe:2012sx}) at a 100 TeV
collider using the monojet channel is $\sim1.4$~TeV.  The reach
can be increased to $\sim3$ TeV using a disappearing tracks search.
Neutralinos up to 1 TeV can be excluded for mass splittings of about 20-30 GeV,
using a soft lepton search \cite{Low:2014cba}. Additionally, Bino-Wino LSPs
with masses up to 1.5 TeV and $\sim 1$~GeV inter-state mass splitting can be
discovered at a 100 TeV collider with $\sim7 \text{ab}^{-1}$ of data.

  \Shufang{Say a few words about typical
  search strategy there.} \Adarsh{Yes, I focused more on well separated spectra
  since that is more similar to our case - should I add more details for the
  0.1 - 50 GeV compressed spectra studies?} \Shufang{Add a couple sentences
  here of the search strategy for the small mass splitting case.}
  \Adarsh{I added a couple of sentences.}

  Well-separated spectra have been studied in \cite{Gori:2014oua,
  Acharya:2014pua} using multi-lepton channels, of the form

\begin{align}
  pp\rightarrow \widetilde{\chi}_i\widetilde{\chi}_j\rightarrow VV +
  \N_1\N_1\rightarrow 2\ell/3\ell + \met,
\end{align}

\noindent where $\widetilde{\chi}_{i,j}$ can be a neutralino or chargino, $\N_1$
is the lightest neutralino, and $V$ being any of the SM gauge bosons $W$, $Z$
or scalar $h$.  These searches work well since the large mass difference
between the electroweakinos can lead to energetic leptons that can be easily
identified. Among the multi-lepton searches for Higgsinos, the trilepton
searches: $\N_{2,3}\C_1\pm\rightarrow WZ\N_1\N_1\rightarrow \ell\ell\ell +
\met$, with $\ell=e, \ \mu$,  have the best reach, due to the relatively high
production cross-section of chargino-neutralino pairs combined with the large
reduction in $t\overline{t}$ and QCD backgrounds obtained by requiring three
leptons.  Higgsino NLSPs can be excluded at 95\% C.L. with mass up to 2.8 TeV
or be discovered with  mass up to 1.8 TeV, for a massless Bino LSP.   The reach
reduces for larger $M_1$, which vanishes for $M_1\gtrsim 600$ GeV (1000 GeV)
for discovery (exclusion).    Multilepton searches with \emph{Zh} and \emph{ZZ}
as the intermediate dibosons are unlikely to be as powerful, due to the smaller
pair-production cross-section of neutral Higgsinos, combined with the low
branching ratio of \emph{Z/h} to leptons. However, if we move beyond
multilepton searches, the channel with \emph{Zh} (originally suggested in
\cite{Han:2013kza}) as the intermediate dibosons emerges as a possible
competitor to the \emph{WZ} channel.

In the left panel of Fig.~\ref{fig:xsection_plot}, we show the pair-production
cross section $\N_2\N_3$ for the Higgsino-like NLSPs.  The cross section varies
between 200 fb to 0.5 fb for Higgsino masses between 500 GeV and 2500 GeV. The
right panel of Fig.~\ref{fig:xsection_plot} shows the decay branching fractions
of $\chi_{2,3}^0\rightarrow Z/h \chi_1^0$, as calculated in SUSY-HIT
\cite{Djouadi:2006bz}.  While one of the Higgsinos mostly decays to $Z$, the
other Higgsino mostly decays to $h$.    Therefore,  for pair produced Higgsinos
$\N_2\N_3$, the \emph{Zh} mode has the greatest branching fraction.  After
multiplying all the branching fractions, the signal cross sections of

\begin{align}
  pp\rightarrow \N_2\N_3\rightarrow (Z\N_1)(h\N_1)\rightarrow bb\ell\ell \N_1\N_1.
\end{align}

\noindent become comparable  to trilepton processes, as shown in table
\ref{tab:xsections}. $Zh+\met$ signal thus provide an alternative discovery
channel for Higgsinos. The signal contains two \emph{b}-jets, two same flavor,
opposite sign leptons, and missing transverse energy. The main backgrounds for
this process are: $t\overline{t}$, \emph{tbW} with the $b$ and $W$ not coming
from a \emph{t}, and \emph{bbWW} with no intermediate $s$-channel top quarks.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.49\textwidth]{images/xsection_plot.pdf}
  \includegraphics[width=0.49\textwidth]{images/br_plot.pdf}
  \caption{Higgsino pair production $\N_2\N_3$
    cross section (left panel), calculated using
    Prospino2~\cite{Beenakker:1999xh},  and decay branching fractions
    (right panel), calculated in SUSY-HIT~\cite{Djouadi:2006bz},  as a function
    of $\mu$ for $M_1=25$ GeV. However, it should be noted that the dependence
    on $M_1$ is very weak.  \Shufang{Could you use $\mu$ for both xlabels?  The
    CS and Br will depend slightly on the sign of mu.  Since we do use positive
  mu in our analyses, let's label it this way.  I have a footnote in intro to
explain the sign effect of mu.  Also, change right panel, ylabel to be either
Br or Branching Fraction.}
}
  \label{fig:xsection_plot}
\end{figure}

\begin{table}
  \centering
  \begin{tabular}{l|rr}
    \toprule
    Stage & $\N_{2,3}\C_1$ & $\N_2\N_3$\\
    \midrule
    Pair production cross section & 60 fb & 17.2 fb\\
    Intermediate diboson contribution & ($WZ$) 29.7 fb  & ($Zh$) 9.15 fb \\
    Applying \emph{BR}$(W\rightarrow \ell\nu)$, \emph{BR}$(Z\rightarrow \ell\ell)$ \& \emph{BR}$(h\rightarrow bb)$ & 0.42 fb & 0.37 fb\\
    \bottomrule
  \end{tabular}
  \caption{Comparison of Higgsino pair production cross-sections for
    $(\N_{2,3}\C_1)$ and
    $(\N_{2}\N_{3})$, for $\mu\approx 1$ TeV, at a 100
    TeV \emph{pp} collider. 
    The pair-production cross section for $\N_{2,3}\C_1$ is taken from
    \cite{Gori:2014oua}, while the pair-production cross-section for $\N_2\N_3$
    is calculated at Next-to-Leading Order (NLO) using Prospino2. The branching
    ratios to the diboson intermediate states are calculated using SUSY-HIT,
    and the branching ratios to the SM final states are taken from
    \cite{Olive:2016xmw}.}
  \label{tab:xsections}
\end{table}
